---
title: "Convergence Concepts"
author: "Xiaoxia Champon & Justin Post"
date: "2023-04-18"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The plots below don't show up where they should but this is how I port my apps in. I use the `shinyApp()` function and then specify the `shinyUI()` within that. From there you can just pull in the UI elements. When you close that function, just put a comma after it and then place the relevant `shinyServer()` code below that. Then you close up the `shinyApp()` function and that's it.

```{r echo=FALSE, message = FALSE}
library(shiny)
library(shinydashboard)
#number of data sets to create
NumData<-10000

shinyApp(
  shinyUI(
      fluidRow(
          column(3,
               selectizeInput("dist",label=h3("Distribution"),selected="Uniform",choices=sort(c("Uniform","Normal","Gamma","Exponential","Chi-Square","Beta","Binomial","Geometric","Poisson","Cauchy"))),
               numericInput("param1","Lower Limit",value=0,step=1),
               conditionalPanel(condition= "(input.dist=='Uniform')||(input.dist=='Normal')||(input.dist=='Gamma')||(input.dist=='Beta')||(input.dist=='Cauchy')",numericInput("param2","Upper Limit",value=1,step=1)), 
               selectizeInput("stat",label=h3("Statistic"),selected="Sample Mean",choices=c("Sample Mean","Standardized Sample Mean","Sample Variance","Sample Standard Deviation","Sample Median","Sample Max","Sample Min")),
               numericInput("n",label=h4("Sample Size (2 if blank)"),value=10,min=2,step=1),
               sliderInput("N", label=h4("Number of Data Sets:"),min = 1, max = 10000, value = 1, step = 1,animate=list(TRUE, interval=350,loop=TRUE)),
               numericInput("rows",label=h4("Number of Rows of Data to Display"),value=5,step=1,min=1,max=25)
          ), 
          column(9,
            fluidRow(
                #Show a plot of the parent population  
                plotOutput("parentPlot"),
              # Show a plot of the sampling distribution
                plotOutput("samplePlot")
            ),
            br(),
            fluidRow(
                plotOutput("statPlot"),
                tableOutput("sampStatTable"),
                tableOutput("statTable"),
                tableOutput("dataTable")   
              )
          )
      )
  ), 
  
  # Define server logic required to draw the plots
  shinyServer(function(input, output,session) {
    
    #generate data
    simData<-reactive({
      #get reactive things and inputs
      input$dist
      input$n
      input$param1
      input$param2
          
      #sample size
      n<-input$n
      #Num samples
      dist<-input$dist
          
      if (is.na(n)){n<-2}
      
      #get samples from appropriate distribution 
      if (dist=="Uniform"){
        min<-input$param1
        max<-input$param2
        samples<-matrix(runif(n*NumData,min=min,max=max),nrow=NumData,ncol=n)
      } else if (dist=="Normal") {
        mean<-input$param1
        sd<-input$param2
        samples<-matrix(rnorm(n*NumData,mean=mean,sd=sd),nrow=NumData,ncol=n)
      } else if (dist=="Gamma") {
        alpha<-input$param1
        lambda<-input$param2
        samples<-matrix(rgamma(n*NumData,shape=alpha,rate=lambda),nrow=NumData,ncol=n)
      } else if (dist=="Exponential") {
        lambda<-input$param1
        samples<-matrix(rexp(n*NumData,rate=lambda),nrow=NumData,ncol=n)
      } else if (dist=="Chi-Square") {
        df<-input$param1
        samples<-matrix(rchisq(n*NumData,df=df),nrow=NumData,ncol=n)
      } else if (dist=="Beta") {
        alpha<-input$param1
        beta<-input$param2
        samples<-matrix(rbeta(n*NumData,shape1=alpha,shape2=beta),nrow=NumData,ncol=n)
      } else if (dist=="Binomial") {
        prob<-input$param1
        samples<-matrix(rbinom(n*NumData,size=1,prob=prob),nrow=NumData,ncol=n)
      } else if (dist=="Geometric") {
        prob<-input$param1
        samples<-matrix(rgeom(n*NumData,prob=prob),nrow=NumData,ncol=n)
      } else if (dist=="Poisson") {
        lambda<-input$param1
        samples<-matrix(rpois(n*NumData,lambda=lambda),nrow=NumData,ncol=n)
      } else if (dist=="Cauchy") {
        location<-input$param1
        scale<-input$param2
        samples<-matrix(rcauchy(n*NumData,location=location,scale=scale),nrow=NumData,ncol=n)
      }
      
      #return data
      samples
    })
  
   
    #update parameters in ui  
    observe({
      #get input
      dist <- input$dist
          
      #update the values of the parameters in the ui based on the distribution
      if (dist=="Uniform"){
        updateNumericInput(session, "param1", label="Lower Limit",value = 0,step=1)
        updateNumericInput(session, "param2", label="Upper Limit",value = 1,step=1)
      } else if (dist=="Normal") {
        updateNumericInput(session, "param1", label="Mean",value = 0,step=1)
        updateNumericInput(session, "param2", label="Standard Deviation",value = 1,min=0,step=1)
      } else if (dist=="Gamma") {
        updateNumericInput(session, "param1", label="Alpha (Shape)",value = 1,min=0,step=1)
        updateNumericInput(session, "param2", label="Lambda (Rate)",value = 1,min=0,step=1)
      } else if (dist=="Exponential") {
        updateNumericInput(session, "param1", label="Lambda (Rate)",value = 1,min=0,step=1)
        updateNumericInput(session, "param2", label="Not Applicable",value = 0,min=0,max=0)
      } else if (dist=="Chi-Square") {
        updateNumericInput(session, "param1", label="DF",value = 1,min=0,step=1)
        updateNumericInput(session, "param2", label="Not Applicable",value = 0,min=0,max=0)
      } else if (dist=="Beta") {
        updateNumericInput(session, "param1", label="alpha",value = 1,min=0,step=1)
        updateNumericInput(session, "param2", label="beta",value = 1,min=0,step=1)
      } else if (dist=="Binomial") {
        updateNumericInput(session, "param1", label="Probability of Success",value = 0.5,min=0,max=1,step=0.05)
        updateNumericInput(session, "param2", label="Not Applicable",value = 0,min=0,max=0)
      } else if (dist=="Geometric") {
        updateNumericInput(session, "param1", label="Probability of Success",value = 0.5,min=0,max=1,step=0.05)
        updateNumericInput(session, "param2", label="Not Applicable",value = 0,min=0,max=0)
      } else if (dist=="Poisson") {
        updateNumericInput(session, "param1", label="Lambda (Mean)",value = 1,min=0,step=1)
        updateNumericInput(session, "param2", label="Not Applicable",value = 0,min=0,max=0)
      } else if (dist=="Cauchy") {
        updateNumericInput(session, "param1", label="Location",value = 0,step=1)
        updateNumericInput(session, "param2", label="Scale",value = 1,min=0,step=1)
      }  
  
    })
  
    
    #create parent population plot
    output$parentPlot<-renderPlot({
      #input distribution
      dist<-input$dist
      #sample size
      n<-input$n
      
      #choose appropriate distribution  
      if (dist=="Uniform"){
        #parameters
        min<-input$param1
        max<-input$param2
        #Plotting sequence
        x <- seq(from=min,to=max,length=2)
        #draw the parent distribution plot
        plot(x=x,y=dunif(x=x,min=min,max=max),main=paste(dist," Density"),
             xlab="y", ylab="f(y)",type="l")
      } else if (dist=="Normal") {
        #parameters
        mean<-input$param1
        sd<-input$param2
        #Plotting sequence
        x <- seq(from=mean-4*sd,to=mean+4*sd,length=1000)
        #draw the parent distribution plot
        plot(x=x,y=dnorm(x=x,mean=mean,sd=sd),main=paste(dist," Density"),
             xlab="y", ylab="f(y)",type="l")
      } else if (dist=="Gamma") {
        #parameters
        alpha<-input$param1
        lambda<-input$param2
        #Plotting sequence
        x <- seq(from=0,to=alpha/lambda+4*sqrt(alpha/lambda^2),length=1000)
        #draw the parent distribution plot
        plot(x=x,y=dgamma(x=x,shape=alpha,rate=lambda),main=paste(dist," Density"), xlab="y", ylab="f(y)",type="l")
      } else if (dist=="Exponential") {
        #parameters
        lambda<-input$param1
        #Plotting sequence
        x <- seq(from=0,to=5/lambda,length=1000)
        #draw the parent distribution plot
        plot(x=x,y=dexp(x=x,rate=lambda),main=paste(dist," Density"),xlab="y", ylab="f(y)",type="l")
      } else if (dist=="Chi-Square") {
        #parameters
        df<-input$param1
        #Plotting sequence
        x <- seq(from=0,to=df+4*sqrt(2*df),length=1000)
        #draw the parent distribution plot
        plot(x=x,y=dchisq(x=x,df=df),main=paste(dist," Density"), xlab="y", ylab="f(y)",type="l")
      } else if (dist=="Beta") {
        #Plotting sequence
        x <- seq(from=0,to=1,length=1000)
        #parameters
        alpha<-input$param1
        beta<-input$param2
        #draw the parent distribution plot
        plot(x=x,y=dbeta(x=x,shape1=alpha,shape2=beta),main=paste(dist," Density"), xlab="y", ylab="f(y)",type="l")
      } else if (dist=="Binomial") {
        #Plotting sequence
        x <- seq(from=0,to=n,by=1)
        #parameters
        prob<-input$param1
        #draw the parent distribution plot
        plot(x=x,y=dbinom(x=x,size=n,prob=prob),main=paste(dist," Density"), xlab="y", ylab="f(y)",type="h")
      } else if (dist=="Geometric") {
        #parameters
        prob<-input$param1
        #Plotting sequence
        x<-seq(from=0,to=1/prob+4*sqrt((1-prob)/prob^2)+2,by=1)
        #draw the parent distribution plot
        plot(x=x,y=dgeom(x=x,prob=prob),main=paste(dist," Density"),xlab="y", ylab="f(y)",type="h")
      } else if (dist=="Poisson") {
        #parameters
        lambda<-input$param1
        #Plotting sequence
        x <- seq(from=0,to=lambda+4*sqrt(lambda),by=1)
        #draw the parent distribution plot
        plot(x=x,y=dpois(x=x,lambda=lambda),main=paste(dist," Density"), xlab="y", ylab="f(y)",type="h")
      } else if (dist=="Cauchy") {
        #Plotting sequence
        #parameters
        location<-input$param1
        scale<-input$param2
        x <- seq(from=location-6*scale,to=location+6*scale,length=1000)
        #draw the parent distribution plot
        plot(x=x,y=dcauchy(x=x,location=location,scale=scale),main=paste(dist," Density"), xlab="y", ylab="f(y)",type="l")
      }
      
    })
    
    
    #get plot of sample data for current data set
    output$samplePlot<-renderPlot({
      #input data set number
      N<-input$N
      #sample size
      n<-input$n
      #Get data for data set N
      samples<-simData()
      samples<-samples[N,]
  
      #Get statistic being used
      statistic<-input$stat
  
      #create histogram and save it, need to find where stat lies so we can color it
      temp<-hist(samples,main=paste("Histogram of Sample ",N,"'s data"),xlab="Data Values, y",prob=TRUE)    
  
      #calculate value of statistic
      if ((statistic=="Sample Mean")||(statistic=="Standardized Sample Mean")){
        statval<-mean(samples)
      } else if (statistic=="Sample Variance"){
        statval<-var(samples)
      } else if (statistic=="Sample Standard Deviation"){
        statval<-sd(samples)
      } else if (statistic=="Sample Median"){
        statval<-median(samples)
      } else if (statistic=="Sample Max"){
        statval<-max(samples)
      } else if (statistic=="Sample Min"){
        statval<-min(samples)
      }
          
      #get the break from the histogram where the stat occurs
      index<-which((temp$breaks[1:(length(temp$breaks)-1)]<statval)*(statval<temp$breaks[2:length(temp$breaks)])==1)
      #create the color vector to color the histogram
      clr<-rep("white",length(temp$breaks))
      if ((statistic=="Sample Mean")||(statistic=="Standardized Sample Mean")||(statistic=="Sample Median")||(statistic=="Sample Max")||(statistic=="Sample Min")){
        clr[index]<-"blue"
      }
        #create histogram
        hist(samples,main=paste("Histogram of Sample ",N,"'s data"),xlab="Data Values, y",prob=TRUE,col=clr)    
        
    })
  
      
    #get histogram of statistic based on data sets so far
    output$statPlot <- renderPlot({
      #sample size
      n<-input$n
      #Num samples
      N<-input$N
      #distribution
      dist<-input$dist
      #stat
      statistic<-input$stat
      
      #figure out what mean and standard deviation of parent are for z-score
      if (dist=="Uniform"){
        min<-input$param1
        max<-input$param2
        meanz<-(min+max)/2
        varz<-(max-min)^2/12
      } else if (dist=="Normal") {
        mean<-input$param1
        sd<-input$param2
        meanz<-mean
        varz<-sd^2
      } else if (dist=="Gamma") {
        alpha<-input$param1
        lambda<-input$param2
        meanz<-alpha/lambda
        varz<-alpha/lambda^2
      } else if (dist=="Exponential") {
        lambda<-input$param1
        meanz<-1/lambda
        varz<-1/lambda^2
      } else if (dist=="Chi-Square") {
        df<-input$param1
        meanz<-df
        varz<-2*df
      } else if (dist=="Beta") {
        alpha<-input$param1
        beta<-input$param2
        meanz<-alpha/(alpha+beta)
        varz<-(alpha*beta)/((alpha+beta)^2*(alpha+beta+1))
      } else if (dist=="Binomial") {
        prob<-input$param1
        meanz<-n*prob
        varz<-n*prob*(1-prob)
      } else if (dist=="Geometric") {
        prob<-input$param1
        meanz<-1/prob
        varz<-(1-prob)/prob^2
      } else if (dist=="Poisson") {
        lambda<-input$param1
        meanz<-varz<-lambda
      } else if (dist=="Cauchy") {
        location<-input$param1
        scale<-input$param2
        meanz<-location
        varz<-scale
      }
      
      #get values of the stat for each sample
      samples<-simData()
      if(N==1){
        if (statistic=="Sample Mean"){
          statvals<-mean(samples[1,])
        } else if (statistic=="Standardized Sample Mean"){
          statvals<-(mean(samples[1,])-meanz)/sqrt(varz/n)
        } else if (statistic=="Sample Variance"){
          statvals<-var(samples[1,])
        } else if (statistic=="Sample Standard Deviation"){
          statvals<-sd(samples[1,])
        } else if (statistic=="Sample Median"){
          statvals<-median(samples[1,])
        } else if (statistic=="Sample Max"){
          statvals<-max(samples[1,])
        } else if (statistic=="Sample Min"){
          statvals<-min(samples[1,])
        }
      } else if (N>1){
        if (statistic=="Sample Mean"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=mean)
        } else if (statistic=="Standardized Sample Mean"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=function(x,meanz,varz){(mean(x)-meanz)/sqrt(varz/n)}, meanz,varz)
        } else if (statistic=="Sample Variance"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=var)
        } else if (statistic=="Sample Standard Deviation"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=sd)
        } else if (statistic=="Sample Median"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=median)
        } else if (statistic=="Sample Max"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=max)
        } else if (statistic=="Sample Min"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=min)
        }
      }
  
      #get the break from the histogram where the stat for data set N occurs
      temp<-hist(statvals,main=paste("Sampling Distribution of the",statistic),prob=TRUE,xlab="Observed Values") 
      index<-which((temp$breaks[1:(length(temp$breaks)-1)]<statvals[N])*(statvals[N]<temp$breaks[2:length(temp$breaks)])==1)
      #create the color vector to color the histogram
      clr<-rep("white",length(temp$breaks))
      clr[index]<-"blue"
      
      # draw the sampling distribution
      if ((dist=="Cauchy")&&(N>1)&&(N<2000)){
        hist(statvals,main=paste("Sampling Distribution of the",statistic),prob=TRUE,xlab="Observed Values",breaks=100,col=clr) 
      } else {
        hist(statvals,main=paste("Sampling Distribution of the",statistic),prob=TRUE,xlab="Observed Values",col=clr) 
      }
  
    })
  
    
    #get table of data
    output$dataTable<-renderTable({
      #input # of rows to display
      rows<-input$rows
      #sample size
      n<-input$n
      rows<-min(n,rows)
      #Number of data set
      N<-input$N
      #Get data for data set N
      samples<-simData()
      datadisplay<-data.frame(Obs=1:rows,DataValue=sprintf("%.4f",samples[N,1:rows]))
      datadisplay
    },include.colnames=FALSE)
  
    
    #get summary stats for sample
    output$statTable<-renderTable({
      #Number of data set
      N<-input$N
      #Get data for data set N
      samples<-simData()
      samples<-samples[N,]
      
      #find stats
      statdisplay<-data.frame(Stats=c("Sample Mean","Sample Var","Sample SD","Sample Min","Sample Median","Sample Max"), Value=c(sprintf("%.4f",mean(samples)),sprintf("%.4f",var(samples)),sprintf("%.4f",sd(samples)), sprintf("%.4f",min(samples)),sprintf("%.4f",median(samples)),sprintf("%.4f",max(samples))))
  
      statdisplay
    },include.colnames=FALSE)
      
    
    #get statistics about the sampling distribution  
    output$sampStatTable<-renderTable({
      #sample size
      n<-input$n
      #Num samples
      N<-input$N
      #distribution
      dist<-input$dist
      #stat
      statistic<-input$stat
      
      #figure out what mean and standard deviation of parent are for z-score
      if (dist=="Uniform"){
        min<-input$param1
        max<-input$param2
        meanz<-(min+max)/2
        varz<-(max-min)^2/12
      } else if (dist=="Normal") {
        mean<-input$param1
        sd<-input$param2
        meanz<-mean
        varz<-sd^2
      } else if (dist=="Gamma") {
        alpha<-input$param1
        lambda<-input$param2
        meanz<-alpha/lambda
        varz<-alpha/lambda^2
      } else if (dist=="Exponential") {
        lambda<-input$param1
        meanz<-1/lambda
        varz<-1/lambda^2
      } else if (dist=="Chi-Square") {
        df<-input$param1
        meanz<-df
        varz<-2*df
      } else if (dist=="Beta") {
        alpha<-input$param1
        beta<-input$param2
        meanz<-alpha/(alpha+beta)
        varz<-(alpha*beta)/((alpha+beta)^2*(alpha+beta+1))
      } else if (dist=="Binomial") {
        prob<-input$param1
        meanz<-n*prob
        varz<-n*prob*(1-prob)
      } else if (dist=="Geometric") {
        prob<-input$param1
        meanz<-1/prob
        varz<-(1-prob)/prob^2
      } else if (dist=="Poisson") {
        lambda<-input$param1
        meanz<-varz<-lambda
      } else if (dist=="Cauchy") {
        location<-input$param1
        scale<-input$param2
        meanz<-location
        varz<-scale
      }
        
      #get values of the stat for each sample
      samples<-simData()
      if(N==1){
        if (statistic=="Sample Mean"){
          statvals<-mean(samples[1,])
        }else if (statistic=="Standardized Sample Mean"){
          statvals<-(mean(samples[1,])-meanz)/sqrt(varz)
        } else if (statistic=="Sample Variance"){
          statvals<-var(samples[1,])
        } else if (statistic=="Sample Standard Deviation"){
          statvals<-sd(samples[1,])
        } else if (statistic=="Sample Median"){
          statvals<-median(samples[1,])
        } else if (statistic=="Sample Max"){
          statvals<-max(samples[1,])
        } else if (statistic=="Sample Min"){
          statvals<-min(samples[1,])
        }
      } else if (N>1){
        if (statistic=="Sample Mean"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=mean)
        } else if (statistic=="Standardized Sample Mean"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=function(x,meanz,varz){(mean(x)-meanz)/sqrt(varz/n)}, meanz,varz)
        } else if (statistic=="Sample Variance"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=var)
        } else if (statistic=="Sample Standard Deviation"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=sd)
        } else if (statistic=="Sample Median"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=median)
        } else if (statistic=="Sample Max"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=max)
        } else if (statistic=="Sample Min"){
          statvals<-apply(X=samples[1:N,],MARGIN=1,FUN=min)
        }
      }
      sampstatdisplay<-data.frame(Stats=c("Sample Mean","Sample Var","Sample SD","Sample Min","Sample Median","Sample Max"),Value=c(sprintf("%.4f",mean(statvals)),sprintf("%.4f",var(statvals)),sprintf("%.4f",sd(statvals)),sprintf("%.4f",min(statvals)),sprintf("%.4f",median(statvals)),sprintf("%.4f",max(statvals))))
      
      sampstatdisplay
    },include.colnames=FALSE)
    
  })    
)
  
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r echo=FALSE, eval = FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



